% +--------------------------------------------------------------------+
% | Acknowledgements Page (Optional)
% +--------------------------------------------------------------------+

\newpage
\vspace*{0.9cm}
\begin{center}
{\bf \Huge Acknowledgments}
\end{center}

\setlength{\baselineskip}{0.8cm}

%\pdfbookmark[0]{Acknowledgements}{PDF_Acknowledgements}

% +--------------------------------------------------------------------+
% | Enter text for your acknowledgements in the space below this box.
% |                                                                    
% +--------------------------------------------------------------------+

% thank my dad, mom, and sister
I would like to begin by thanking my father, John, who taught me about toughness, humility, and responsibility - a set of values that proved more useful in grad school than the most complex of equations.
I thank my mother, Donna, who continually pushed me to challenge my intellect as a child and showed me the value of critical thinking.
I thank my sister, Brooke, whose constant encouragement and friendship inspired me to strive to do my best while remaining at K-State.
% thank chi rho
I would also like to thank my roommates and brothers at the Catholic Household of Chi Rho for providing an invaluable support system through my time in grad school.
% thank my fellow grad students
I greatly appreciate all of the help, teamwork, and friendship from my fellow graduate students at KSU - Dan, Richard, Rabab, Kevin, Ye, and Mike.
% thank my undergrads
I would like to thank two undergraduate researchers, Keith and Tyler, for their company and contributions to the work presented in this thesis.
% thank my comittee
I thank Dr. Amir Bahadori and Dr. William Dunn for being on my committee.
% thank doctor roberts
And finally, I would like to thank Dr. Jeremy Roberts for taking me on in his research group and offering continued guidance and encouragment.
I could have never guessed five years ago that his fascinating presentation of nuclear physics would have inspired this academic adventure of mine.

